\bfseries\large Phosphatidylinositol 4,5-bisphosphate regulates cilium transition zone maturation in Drosophila

### 2018-02-08 latex/paper.tex
updated

\documentclass[12pt, twoside, letterpaper]{article}
\usepackage[margin=0.7in]{geometry}
\usepackage[parfill]{parskip}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{authblk}
\usepackage{charter}
\usepackage[numbers]{natbib}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{titlesec}

@@

\title{\bfseries\large Phosphatidylinositol 4,5-bisphosphate regulates cilium transition zone maturation in Drosophila}
\author[1,2]{Alind Gupta}
\author[1,2]{Julie A. Brill}
\affil[1]{Department of Molecular Genetics, University of Toronto, Toronto, ON M5S 1A8, Canada}
\affil[2]{Program in Cell Biology, Hospital for Sick Children, Toronto, ON M5G 1X8, Canada}
\date{}

@@

In this study, we show that the lipid \PIP{} is essential for proper
transition zone maturation in the Drosophila male germline.
Reduction of cellular \PIP{} by expression of the bacterial PIP
phosphatase SigD or inactivation of the type I PIP kinase Skittles
causes developing TZs to elongate excessively.

We had previously shown that reduction of \PIP{} by SigD expression
affected flagellar biogenesis and axoneme structure
\citep{wei2008depletion}.
We also showed aberrant ``comet-shaped'' localization of Unc that
failed to assemble TZ-associated puncta in some cases.
Here, we show that this occurs prior to meiosis,
is not an artefact due to cytokinesis defects
and do not represent an inability of the axoneme to grow.

The type I phosphatidylinositol 4-phosphate kinase (PIPKI) Skittles
suppresses TZ elongation in \sigd{} cells in a cilium-autonomous
manner.
We also found that cells mutant for \textit{sktl} contained hyperelongated
TZs similar to SigD-expressing cells.
Sktl also localizes to centrioles.

It has previously been suggested that LCCS3 might represent a ciliopathy.
PIPKI$\gamma$ is linked to LCCS3 with many features reminiscent
of ciliopathies, such as micrognathia, and a role for the LCCS1 protein
Gle3 in cilia.
In addition, all three human PIPKI isoforms are highly expressed in testes
and PIPKI$\alpha$ and PIPKI$\beta$ are important for fertility in mice
\citep{hasegawa2012phosphatidylinositol}.

Members of the exocyst complex, such as Sec10 and Sec6, have been shown
to be important for cilium formation in cultured cell lines and zebrafish
\citep{zuo2009exocyst, lobo2017exocyst, seixas2016arl13b}.
A mutation in \textit{EXOC8} has been linked
to Joubert syndrome \citep{dixon2012exome}.
The exocyst is known to bind to \PIP{}, and we previously described that
the \textit{onr} allele of Drosophila \textit{Exo84} phenocopied
male germline defects observed in \sigd{}.

@@

\subsection{Statistical methods}
Statistical analysis and graphing was performed using R.
Statistical tests for ``absence of phenotype'' were computed under a
hypothesis assuming binomial likelihood.
All $t$-tests were unpaired and two-sided.
A significance level of 0.01 was fixed \textit{a priori} for all analysis.

@@

\section{Acknowledgements}
We thank Brian Ciruna for input on results.

>>>>

\documentclass[12pt, twoside, letterpaper]{article}
\usepackage[margin=0.6in]{geometry}
\usepackage[parfill]{parskip}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{authblk}
\usepackage{charter}
\usepackage[numbers]{natbib}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{titlesec}

@@

\title{\bfseries\large Phosphatidylinositol 4,5-bisphosphate regulates cilium transition zone maturation in Drosophila}
\author[1,2]{Alind Gupta \thanks{\url{alind.gupta@mail.utoronto.ca}}}
\author[1,2]{Julie A. Brill \thanks{\url{julie.brill@sickkids.ca} (Corresponding author)}}
\affil[1]{Department of Molecular Genetics, University of Toronto, Toronto, ON M5S 1A8, Canada}
\affil[2]{Program in Cell Biology, Hospital for Sick Children, Toronto, ON M5G 1X8, Canada}
\date{}

@@

Cilia are essential for proper human development, and their disruption
can cause genetic disorders which display severity ranging from congenital
lethality to adult onset.
A large number of ciliopathies are associated with genes that encode for
proteins that function at the cilium transition zone (TZ).
Often, a single gene is linked to multiple ciliopathies with variable phenotypes
and severities.
For example, the conserved TZ protein CEP290 is linked to at least six different
cilium-associated disorders, suggesting that many of these encode for mutant proteins
that are hypomorphic in nature whose function may
be affected by modifier mutations and genetic environment.
Animal models with mild perturbations of TZ function may therefore provide
insight on intermediate phenotypes associated with human disease.
Here, we found that reduced levels of the membrane lipid \PIP{},
which is known to be enriched at the base of cilia,
affects TZ maturation, but not assembly,
leading to hyperelongated TZs that show signs of functional defects, and
may represent one such modifier for TZ function.

In the Drosophila male germline, TZ maturation follows the initial
assembly of TZ proteins on the nascent basal body in the cytoplasm
followed by anchoring at the plasma membrane.
TZ maturation involves microtubule rearrangements and TZ elongation,
which leads to cilium-membrane attachment.
Loss of \PIP{}, a membrane lipid largely at the plasma membrane,
disrupts this process.
It causes excessive lengthening of the TZ,
and disrupts membrane-cilium connections.

The type I phosphatidylinositol 4-phosphate kinase (PIPKI) Skittles
suppresses TZ elongation in \sigd{} cells in a cilium-autonomous
manner.
We also found that cells mutant for \textit{sktl} contained hyperelongated
TZs similar to SigD-expressing cells.
Sktl also localizes to centrioles,
and our results suggest that Skittles act locally at cilia to
regulate TZ length.
The protein CEP135, human ortholog of the Drosophila protein
Asterless, has been show to bind PIPKI$\gamma$ in cell lines.
PIPKI$\gamma$ is linked to lethal congenital contractural syndrome
type 3 (LCCS3).
It has previously been suggested that LCCS3 might represent a ciliopathy.
PIPKI$\gamma$ is linked to LCCS3 with many features reminiscent
of ciliopathies, such as micrognathia, and a role for the LCCS1 protein
Gle3 in cilia.
In addition, all three human PIPKI isoforms are highly expressed in testes
and PIPKI$\alpha$ and PIPKI$\beta$ are important for fertility in mice
\citep{hasegawa2012phosphatidylinositol}.

Members of the exocyst complex, such as Sec10 and Sec6, have been shown
to be important for cilium formation in cultured cell lines and zebrafish
\citep{zuo2009exocyst, lobo2017exocyst, seixas2016arl13b}.
The exocyst is known to bind to \PIP{}, and we previously described that
the \textit{onr} allele of Drosophila \textit{Exo84} phenocopied
male germline defects observed in \sigd{}.
In this paper, we show that the \textit{onr} allele of the Drosophila Exo84
phenocopies the loss of cilium-membrane contacts in \sigd{},
but not TZ hyperelongation.
Proper TZ function is required for membrane tethering, as \textit{dila; cby}
mutants, which contain abnormal TZs also show cytoplasmic cilia similar to
\sigd{}.
Our results suggest that the exocyst is required after the initial stages
of TZ assembly and maturation for TZ function.

@@

\subsection{Statistical methods}
Statistical analysis and graphing was performed using R.
A Gaussian jitter was applied when plotting
results in Figures 1 and 2,
but raw data was used for all analysis.
Statistical tests for ``absence of phenotype'' were computed under a
hypothesis assuming binomial likelihood (i.e. assuming a fixed probability of the
phenotype occurring).
All $t$-tests were unpaired and two-sided.
A significance level of 0.01 was fixed \textit{a priori} for all classical analysis.
Bayesian analysis was performed in Stan.
All raw data and code for analysis and plotting can be found online
at \url{http://www.github.com/alindgupta/germline-paper/}.

\subsection{Phylogenetic analysis}
Candidate orthologs of \textit{sktl} or \textit{pip5k59b} were queried from Inparanoid and FlyBase.
Poorly annotated protein sequences were manually curated to ensure
these represented type I phosphatidylinositol phosphate
kinases using reciprocal BLAST search.
Phylogeny.fr (\url{http://www.phylogeny.fr}) was used for
phylogenetic reconstruction with T-Coffee for multiple alignment
and MrBayes for tree construction.
The output was converted to a vector image in Illustrator
and colours were added for the purposes of illustration.

@@

\section{Acknowledgements}
We thank Brian Ciruna for his input on the project
and B. Durand for graciously providing unpublished fly stocks.
